### 2010-03-18 papers/icfp-2010-desc/paper_idesc.tex
Desc@ICFP: Some minor fixes and notes

However, this new class of objects cannot be defined by mere
induction. In the case of vectors, for instance, we have to define the
whole \emph{family} of vector in one go: vectors of all size needs to
be defined at the same time. Where simply-typed languages have
inductive types, we adopt inductive family of types as our basic
grammar of data-types. To this end, we rely on \emph{indexing}. We are
going to transform the $\Desc$ universe into $\IDesc$, the universe of
indexed descriptions. We express inductive families by the $I \To
\IDesc{I}$ type.

@@

In the alternative presentation, we go further and present all our
codes as quotations of standard type-formers. This presentation is
shown in Figure~\ref{fig:type-former-desc}.

@@

\[
\begin{array}{ll}
\stk{
\data \Desc : \Set \where \\
\;\;\begin{array}{@{}l@{\::\:\:}l@{\quad}l}
    \DId            & \Desc                                   \\
    \DUnit          & \Desc                                   \\
    \DProd{}{}      & \PI{\V{D}, \V{D'}}{\Desc} \Desc         \\
    \DSigma         & \PI{\V{S}}{\Set} \PIS{S \To \Desc} \Desc \\
    \DPi            & \PI{\V{S}}{\Set} \PIS{S \To \Desc} \Desc 
\end{array}
}
&
\stk{
\descop{\_\:}{} : \Desc \To \Set \To \Set \\
\begin{array}{@{}l@{\:=\:\:}ll}
\descop{\DId}{X}          &  X                                           \\
\descop{\DUnit}{X}        &  \Unit                                       \\
\descop{\DProd{D}{D'}}{X} &  \TIMES{\descop{D}{X}}{\descop{D'}{X}}       \\
\descop{\DSigma{S}{D}}{X} &  \SIGMA{\V{s}}{S} \descop{D\: s}{X}                \\
\descop{\DPi{S}{D}}{X}    &  \PI{\V{s}}{S} \descop{D\: s}{X}            
\end{array}
}
\end{array}
\]

@@

In the previous section, we have presented the $\Desc$ universe as the
grammar of functors in the category $\Set$. We have seen how to encode
inductive types in this setting. To encode an inductive family indexed
by $\Bhab{\V{I}}{\Set}$, we rely on functors in the category
$\Set^I$. We call these \emph{indexed functors}. In the following, we
implement $I \To \IDesc{I}$ as a grammar for indexed functors:

@@

This generalisation can be achieved by a minor evolution of $\Desc$
(Fig.~\ref{fig:type-former-desc}). Indeed, the functorial nature of $\Desc$
is entirely captured by the $\DId$ code. We get an indexed functor
by turning $\DId$ into an index-dependent $\DVar$ code:

\[\stk{
\data \IDesc{} (\Bhab{\V{I}}{\Set}) : \Set \where \\
\;\;\begin{array}{@{}l@{\::\:\:}l@{\quad}l}
    \DVar{}         & I \To \IDesc{I}                                   \\
    \ldots          & \ldots
\end{array} \\
\\
\idescop{\_\:}{}{} : \PI{\V{I}}{\Set} \IDesc{I} \To (\V{I} \To \Set) \To \Set        \\
\begin{array}{@{}l@{\:=\:\:}ll}
\idescop{\DVar{i}}{I}{P}      &  P~i                                                 \\
\ldots                        &  \ldots
\end{array}
}\]

@@

\[
\begin{array}{ll}
\stk{
\data \IDesc{} (\Bhab{\V{I}}{\Set}) : \Set \where \\
\;\;\begin{array}{@{}l@{\::\:\:}l@{\quad}l}
    \DVar{}         & I \To \IDesc{I}                                   \\
    \DConst{}       & \Set \To \IDesc{I}                                \\
    \DProd{}{}      & \PI{\V{D}, \V{D'}}{\IDesc{I}} \IDesc{I}           \\
    \DSigma         & \PI{\V{S}}{\Set} \PIS{S \To \IDesc{I}} \IDesc{I}  \\
    \DPi            & \PI{\V{S}}{\Set} \PIS{S \To \IDesc{I}} \IDesc{I} 
\end{array}
}
&
\stk{
\idescop{\_\:}{}{} : \PI{\V{I}}{\Set} \IDesc{I} \To (\V{I} \To \Set) \To \Set                  \\
\begin{array}{@{}l@{\:=\:\:}ll}
\idescop{\DVar{i}}{I}{P}      &  P~i                                                 \\
\idescop{\DConst{X}}{I}{P}    &  X                                                   \\
\idescop{\DProd{D}{D'}}{I}{P} &  \TIMES{\idescop{D}{I}{P}}{\idescop{D'}{I}{P}}       \\
\idescop{\DSigma{S}{D}}{I}{P} &  \SIGMA{\V{s}}{S} \idescop{D\: s}{I}{P}                    \\
\idescop{\DPi{S}{D}}{I}{P}    &  \PI{\V{s}}{S} \idescop{D\: s}{I}{P}            
\end{array}
}
\end{array}
\]

@@

In the case of vectors, for instance, for the index $\NatZero$, we
would only propose the constructor $\ListNil$. Similarly, for
$\NatSuc{n}$, we only propose the constructor $\ListCons{}{}$. We will
see more instances of this pattern in the following.

@@

\[\stk{
\IDescD : \PI{\V{I}}{\Set} \IDesc{\Unit} \\
\begin{array}{@{}ll}
\IDescD~I \mapsto \DSigma{}{} & (\EnumT{[ \DVar{},
                                          \DConst{},
                                          \DProd{}{},
                                          \DSigma{}{}, 
                                          \DPi{}{} ]}) \\
                              & \left[\begin{array}{l}
                                      \DConst{I}                  \\
                                      \DConst{\Set}               \\
                                      \DProd{\DVar{\Void}}{\DVar{\Void}}  \\
                                      \DSigma{\Set}{(\LAM{S} \DPi{S}{(\LAM{\_} \DVar{\Void})})} \\
                                      \DSigma{\Set}{(\LAM{S} \DPi{S}{(\LAM{\_} \DVar{\Void})})}
                                   \end{array}
                             \right]
\end{array}
}\]

Note that this description also lives in $\IDesc{\Unit}$. Indeed,
$\IDesc$ is merely an inductive type. As for the $\Desc$ levitation,
the finite function space $[ \ldots ]$ elaborates into a special
purpose $\F{switchID}$ operator.

@@

The notation $\Epolym$ represents a form of ``polymorphism'', meaning
that $\alpha$ can be any type. \note{Polymorph? Will any ghost try to
  haunt me with that?  A ghost under a mixed prefix?
  Aaaaaaaaaaargh\ldots}

@@

\[\stkc{
\stk{
\HExprD : \TagIDesc{\Ty} \\
\HExprD \mapsto ( \HExprAD , \HExprID ) \\
} \\
\\
\begin{array}{ll}
\stk{
\HExprAD : \ATagIDesc{\Ty} \\
\HExprAD \mapsto \left[
                 \begin{array}{l}
                   \EnumT{[\Eval{}, \Econd{}{}{}]} \red{,} \\
                   \LAM{ty}
                   \left[ %%! Should be read
                   \begin{array}{l}
                   \DConst{(\Val{ty})} \\
                   \DProd{\DVar{\Ebool}}{\DProd{\DVar{ty}}{\DVar{ty}}}
                   \end{array}
                   \right]
                 \end{array}
                 \right] %%! Should be \red
\\
\\
} &
\stk{
\HExprID : \ITagIDesc{\Ty} \\
\HExprID \mapsto \left[
                 \begin{array}{l}
                   \LAM{ty} \case{ty}{\Enat \To \EnumT{[\Eplus{}{}]} \\ \Ebool \To \EnumT{[\Ele{}{}]}} \red{,} \\
                   \LAM{ty}
                   \case{ty}{
                     \Enat \To \DProd{\DVar{nat}}{\DVar{nat}} \\
                     \Ebool \To \DProd{\DVar{nat}}{\DVar{nat}}
                   }
                   \end{array}
                   \right]
}
\end{array}
}\]

@@

\note{the big square brackets should be red}

@@

\[\stk{
\evalOne : \PI{\V{ty}}{\Ty} \idescop{\toIDesc{\HExprD}}{\Ty}{ty} \To {\Val{ty}} \\
\begin{array}{@{}l@{\:=\:\:}l}
\evalOne\: \_\: \pair{\Eval{}}{x}{}                                         & x \\
\evalOne\: \_\: \pair{\Econd{}{}{}}{\pair{\BoolTrue}{\pair{x}{\_}{}}{}}{}   & x \\
\evalOne\: \_\: \pair{\Econd{}{}{}}{\pair{\BoolFalse}{\pair{\_}{y}{}}{}}{}  & x \\
\evalOne\: \Enat\: \pair{\Eplus{}{}}{\pair{x}{y}{}}{}                       & \F{plusHost}\: x y \\
\evalOne\: \Ebool\: \pair{\Ele{}{}}{\pair{x}{y}{}}{}                        & \F{leHost}\: x y 
\end{array} \\
\\
\evalH : \PI{\V{ty}}{\Ty} 
           \IMu{\Ty}{\toIDesc{\HExprD}}{ty} \To
           \Val{ty} \\
\evalH\: ty\: term = \F{cataI}\: \Ty\: 
                                 \toIDesc{\HExprD}\: 
                                 \Val{}\: 
                                 \evalOne\: 
                                 ty\: 
                                 term
}\]

@@

\[\stk{
\FinD : \Nat \To \IDesc{\Nat} \\
\begin{array}{@{}ll}
\FinD\: n = \DSigma{}{} & (\EnumT{[ \FinZero , \FinSuc{} ]}) \\
                        & \left[
                          \begin{array}{l}
                            \DSigma{\Nat}{\LAM{m} \DConst{(n \PropEq \NatSuc{m})}} \\
                            \DSigma{\Nat}{\LAM{m} \DProd{\DVar{m}}{\DConst{(n \PropEq \NatSuc{m})}}}
                          \end{array}
                          \right]
\end{array}
}\]

Intuitively, the constraint is captured as follow. First, we store an
element $m$ of $\Nat$, thanks to a $\DSigma$. However, the constraint
stipulates that $m$ cannot be \emph{any} natural numbers: it must be
``the index minus one''. This translates into the constraint $n
\PropEq \NatSuc{m}$.

@@

\note{ I suspect we don't implement vectors, space-wise }

@@

While we support the same grammar, indexed descriptions embraces a
much wider class of data-types. Indeed, we pose no constraint apart
from strict positivity. This suggests that, based on the familiar
grammar of GADTs, we can simply lift off the restrictions and
translates the user's input to a code in our universe of indexed
description.

@@

\[\stk{
\FinD : \Nat \To \IDesc{\Nat} \\
\begin{array}{@{}llll}
\FinD\: 0            & = & \multicolumn{2}{l}{\DSigma{\Void}{\Void} } \\
\FinD\: (\NatSuc{n}) & = & \DSigma{}{} & (\EnumT{[ \FinZero , \FinSuc{} ]}) \\
                     &   &             & \left[
                                         \begin{array}{l}
                                         \DConst{\Unit} \\
                                         \DVar{n}
                                         \end{array}
                                         \right]
\end{array}
}\]

@@

In this section, we develop more structure for our indexed
data-types. In Section~\ref{sec:desc-free-monad}, we have constructed
a free monad operation on simple descriptions. Building on this
experience, we are going to present its equivalent in the indexed
world. The processus is the same. Namely, given an indexed functor, we
derive the indexed functor coding its free monad:

@@

\[\stk{
\HExprAFreeD : \ATagIDesc{\Ty} \\
\HExprAFreeD \mapsto \left[
                 \begin{array}{l}
                   \EnumT{[\Econd{}{}{}]} \red{,} \\
                   \LAM{ty}
                   \left[ %%! Should be red
                   \begin{array}{l}
                   \DProd{\DVar{\Ebool}}{\DProd{\DVar{ty}}{\DVar{ty}}}
                   \end{array}
                   \right]
                 \end{array}
                 \right] %%! Should be \red
}\]

@@

\[\stk{
\IDescFreeD : \PI{\V{I}}{\Set} \IDesc{\Unit} \\
\begin{array}{@{}ll}
\IDescFreeD~I \mapsto \DSigma{}{} & (\EnumT{[ \DConst{},
                                              \DProd{}{},
                                              \DSigma{}{}, 
                                              \DPi{}{} ]}) \\
                                  & \left[\begin{array}{l}
                                        \DConst{\Set}               \\
                                        \DProd{\DVar{\Void}}{\DVar{\Void}}  \\
                                        \DSigma{\Set}{(\LAM{S} \DPi{S}{(\LAM{\_} \DVar{\Void})})} \\
                                        \DSigma{\Set}{(\LAM{S} \DPi{S}{(\LAM{\_} \DVar{\Void})})}
                                    \end{array}\right]
\end{array}
}\]

@@

\[\stk{
\IDescD : \PI{\V{I}}{\Set} \IDesc{\Unit} \\
\IDescD\: I \mapsto \FreeIMonad{[\IDescFreeD]}{\Unit}{I}\: I
}\]

>>>>

However, this new class of objects cannot be defined by mere
induction. In the case of vectors, for instance, we have to define the
whole \emph{family} of vector in one go: vectors of all size needs to
be defined at the same time. Where simply-typed languages have
inductive types, in dependently typed programming we the basic grammar of
data-types should be that of inductive families of types. To this end, we
rely on \emph{indexing}. We are going to transform the $\Desc$ universe into
$\IDesc$, the universe of indexed descriptions. We express inductive families
by the $I \To \IDesc{I}$ type.  \note{pwm: Vec might not be the best
motivating example here, since it doesn't appear in the paper (more on that
later). Come to think of it, this is entirely too vague without spelling out
a concrete example.}

@@

In the alternative presentation, we go further and present all our
codes as quotations of standard type-formers. This presentation is
shown in Figure~\ref{fig:type-former-desc}.
\note{pwm: The exception being non dependent product -- we might want to
explain why this is.}

@@

\begin{figure}

\[\stk{
\begin{array}{ll}
\stk{
\data \Desc : \Set \where \\
\;\;\begin{array}{@{}l@{\::\:\:}l@{\quad}l}
    \DId            & \Desc                                   \\
    \DUnit          & \Desc                                   \\
    \DProd{}{}      & \PI{\V{D}, \V{D'}}{\Desc} \Desc         \\
    \DSigma         & \PI{\V{S}}{\Set} \PIS{S \To \Desc} \Desc \\
    \DPi            & \PI{\V{S}}{\Set} \PIS{S \To \Desc} \Desc 
\end{array}
}
\vspace{0.2in}
\\
\stk{
\descop{\_\:}{} : \Desc \To \Set \To \Set \\
\begin{array}{@{}l@{\:=\:\:}ll}
\descop{\DId}{X}          &  X                                           \\
\descop{\DUnit}{X}        &  \Unit                                       \\
\descop{\DProd{D}{D'}}{X} &  \TIMES{\descop{D}{X}}{\descop{D'}{X}}       \\
\descop{\DSigma{S}{D}}{X} &  \SIGMA{\V{s}}{S} \descop{D\: s}{X}                \\
\descop{\DPi{S}{D}}{X}    &  \PI{\V{s}}{S} \descop{D\: s}{X}            
\end{array}
}
\end{array}
}\]

@@

In the previous section, we have presented the $\Desc$ universe as the
grammar of functors in the category $\Set$. We have seen how to encode
inductive types in this setting. To encode an inductive family indexed
by $\Bhab{\V{I}}{\Set}$, we rely on functors in the category
$\Set^I$. We call these \emph{indexed functors}. 
\note{pwm: What's our story \emph{vis-\`{a}-vis} the 2 notions of indexed
functor at play here: $(I\to\Set)\to\Set$ and $(I\to\Set)\to(I\to\Set)$?}  
In the following, we
implement $I \To \IDesc{I}$ as a grammar for indexed functors:

@@

This generalisation can be achieved by a minor evolution of $\Desc$
(Fig.~\ref{fig:type-former-desc}). Indeed, the functorial nature of the
atomic $\Desc$ is entirely captured by the $\DId$ code. We get an indexed
functor by turning $\DId$ into an index-dependent $\DVar$ code:

\[\stk{
\data \IDesc{} (\Bhab{\V{I}}{\Set}) : \Set \where \\
\;\;\begin{array}{@{}l@{\::\:\:}l@{\quad}l}
    \DVar{}         & I \To \IDesc{I}                                   \\
    \ldots          & \ldots
\end{array} \\
\\
\idescop{\_\:}{}{} :_{\PI{\V{I}}{\Set}} \IDesc{I} \To (\V{I} \To \Set) \To \Set        \\
\begin{array}{@{}l@{\:=\:\:}ll}
\idescop{\DVar{i}}{I}{P}      &  P~i                                                 \\
\ldots                        &  \ldots
\end{array}
}\]

@@

\begin{figure}

\[\stk{
\begin{array}{ll}
\stk{
\data \IDesc{} (\Bhab{\V{I}}{\Set}) : \Set \where \\
\;\;\begin{array}{@{}l@{\::\:\:}l@{\quad}l}
    \DVar{}         & I \To \IDesc{I}                                   \\
    \DConst{}       & \Set \To \IDesc{I}                                \\
    \DProd{}{}      & \PI{\V{D}, \V{D'}}{\IDesc{I}} \IDesc{I}           \\
    \DSigma         & \PI{\V{S}}{\Set} \PIS{S \To \IDesc{I}} \IDesc{I}  \\
    \DPi            & \PI{\V{S}}{\Set} \PIS{S \To \IDesc{I}} \IDesc{I} 
\end{array}
}
\vspace{0.2in}
\\
\stk{
\idescop{\_\:}{}{} :_{\PI{\V{I}}{\Set}} \IDesc{I} \To (\V{I} \To \Set) \To \Set                  \\
\begin{array}{@{}l@{\:=\:\:}ll}
\idescop{\DVar{i}}{I}{P}      &  P~i                                                 \\
\idescop{\DConst{X}}{I}{P}    &  X                                                   \\
\idescop{\DProd{D}{D'}}{I}{P} &  \TIMES{\idescop{D}{I}{P}}{\idescop{D'}{I}{P}}       \\
\idescop{\DSigma{S}{D}}{I}{P} &  \SIGMA{\V{s}}{S} \idescop{D\: s}{I}{P}                    \\
\idescop{\DPi{S}{D}}{I}{P}    &  \PI{\V{s}}{S} \idescop{D\: s}{I}{P}            
\end{array}
}
\end{array}
}\]

@@

In the case of vectors, for instance, for the index $\NatZero$, we
would only propose the constructor $\ListNil$. Similarly, for
$\NatSuc{n}$, we only propose the constructor $\ListCons{}{}$. We will
see more instances of this pattern in the following.
\note{We're talking about vectors again, are we \emph{really} not going to
define them?}

@@

\[\stk{
\IDescD : \PI{\V{I}}{\Set} \IDesc{\Unit} \\
\begin{array}{@{}ll}
\IDescD~I \mapsto \DSigma{}{} & (\EnumT \red{[} \DVar{},
                                          \DConst{},
                                          \DProd{}{},
                                          \DSigma{}{}, 
                                          \DPi{}{} \red{]}) \\
                              & \bigRedBracket{\begin{array}{l}
                                      \DConst{I}                  \\
                                      \DConst{\Set}               \\
                                      \DProd{\DVar{\Void}}{\DVar{\Void}}  \\
                                      \DSigma{\Set}{(\LAM{S} \DPi{S}{(\LAM{\_} \DVar{\Void})})} \\
                                      \DSigma{\Set}{(\LAM{S} \DPi{S}{(\LAM{\_} \DVar{\Void})})}
                                   \end{array}
                             }
\end{array}
}\]

Note that this description also lives in $\IDesc{\Unit}$. Indeed,
for any $\V{I}$, $\IDesc~\V{I}$ is merely an inductive type. As for the
$\Desc$ levitation, the finite function space $[ \ldots ]$ elaborates into a
special purpose $\F{switchID}$ operator.

@@

The notation $\Epolym$ represents a form of ``polymorphism'', meaning
that $\alpha$ can be any type. \note{Polymorph? Will any ghost try to
  haunt me with that?  A ghost under a mixed prefix?
  Aaaaaaaaaaargh\ldots (Wha? - pwm)}
\note{pwm: I don't really get the need for this $\Epolym$ notation. Why not
just abstract over an elem of $\Ty£$?}

@@

\[\stkc{
\stk{
\HExprD : \TagIDesc{\Ty} \\
\HExprD \mapsto ( \HExprAD , \HExprID ) \\
} \\
\\
\begin{array}{ll}
\stk{
\HExprAD : \ATagIDesc{\Ty} \\
\HExprAD \mapsto \bigRedBracket{
                 \begin{array}{l}
                   \EnumT{[\Eval{}, \Econd{}{}{}]} \red{,} \\
                   \LAM{ty}
                   \bigRedBracket{
                   \begin{array}{l}
                   \DConst{(\Val{ty})} \\
                   \DProd{\DVar{\Ebool}}{\DProd{\DVar{ty}}{\DVar{ty}}}
                   \end{array}
                   }
                 \end{array}
                 }
\\
\\
} &
\stk{
\HExprID : \ITagIDesc{\Ty} \\
\HExprID \mapsto \bigRedBracket{
                 \begin{array}{l}
                   \LAM{ty} \case{ty}{\Enat \To \EnumT{[\Eplus{}{}]} \\ \Ebool \To \EnumT{[\Ele{}{}]}} \red{,} \\
                   \LAM{ty}
                   \case{ty}{
                     \Enat \To \DProd{\DVar{nat}}{\DVar{nat}} \\
                     \Ebool \To \DProd{\DVar{nat}}{\DVar{nat}}
                   }
                   \end{array}
                   }
}
\end{array}
}\]

@@

\note{the big square brackets should be red (pwm: use \\bigRedBracket :)) }

@@

\[\stk{
\evalOne : \PI{\V{ty}}{\Ty} \idescop{\toIDesc{\HExprD}}{\Ty}{ty} \To {\Val{ty}} \\
\begin{array}{@{}l@{\:=\:\:}l}
\evalOne\: \_\: \pair{\Eval{}}{x}{}                                         & x \\
\evalOne\: \_\: \pair{\Econd{}{}{}}{\pair{\BoolTrue}{\pair{x}{\_}{}}{}}{}   & x \\
\evalOne\: \_\: \pair{\Econd{}{}{}}{\pair{\BoolFalse}{\pair{\_}{y}{}}{}}{}  & x \\
\evalOne\: \Enat\: \pair{\Eplus{}{}}{\pair{x}{y}{}}{}                       & \F{plusHost}\: x y \\
\evalOne\: \Ebool\: \pair{\Ele{}{}}{\pair{x}{y}{}}{}                        & \F{leHost}\: x y 
\end{array} \\
\\
\evalH : \PI{\V{ty}}{\Ty} 
           \IMu{\Ty}{\toIDesc{\HExprD}}{ty} \To
           \Val{ty} \\
\evalH\: ty\: term = \F{cataI}\: \Ty\: 
                                 \toIDesc{\HExprD}\: 
                                 \Val{}\: 
                                 \evalOne\: 
                                 ty\: 
                                 term
}\]
\note{pwm: $\evalOne$ rather illustrates my point about the 2 types of IFunc
floating around, $\idescop{\_}{}{}$ is defined using the 1st notion, and used it
here as if it were the 2nd}

@@

\[\stk{
\FinD : \Nat \To \IDesc{\Nat} \\
\begin{array}{@{}ll}
\FinD\: n = \DSigma{}{} & (\EnumT{[ \FinZero , \FinSuc{} ]}) \\
                        & \bigRedBracket{
                          \begin{array}{l}
                            \DSigma{\Nat}{\LAM{m} \DConst{(n \PropEq \NatSuc{m})}} \\
                            \DSigma{\Nat}{\LAM{m} \DProd{\DVar{m}}{\DConst{(n \PropEq \NatSuc{m})}}}
                          \end{array}
                          }
\end{array}
}\]

Intuitively, the constraint is captured as follows. First, we store an
element $m$ of $\Nat$, thanks to a $\DSigma$. However, the constraint
stipulates that $m$ cannot be \emph{any} natural numbers: it must be
``the index minus one''. This translates into the constraint $n
\PropEq \NatSuc{m}$.

@@

\note{ I suspect we don't implement vectors, space-wise }
\note{pwm: I suspect we don't get away with leaving them out}

@@

While we support the same grammar, indexed descriptions embraces a
much wider class of data-types. Indeed, we pose no constraint apart
from strict positivity\note{pwm: Is this actually the 1st mention of this
phrase? I think that should be rectified :)}. This suggests that, based on
the familiar grammar of GADTs, we can simply lift off the restrictions and
translates the user's input to a code in our universe of indexed description.

@@

\[\stk{
\FinD : \Nat \To \IDesc{\Nat} \\
\begin{array}{@{}llll}
\FinD\: 0            & = & \multicolumn{2}{l}{\DSigma{\Void}{\Void} } \\
\FinD\: (\NatSuc{n}) & = & \DSigma{}{} & (\EnumT{[ \FinZero , \FinSuc{} ]}) \\
                     &   &             & \bigRedBracket{
                                         \begin{array}{l}
                                         \DConst{\Unit} \\
                                         \DVar{n}
                                         \end{array}
                                         }
\end{array}
}\]

@@

In this section, we develop more structure for our indexed
data-types. In Section~\ref{sec:desc-free-monad}, we have constructed
a free monad operation on simple descriptions. Building on this
experience, we are going to present its equivalent in the indexed
world. The process the same. Namely, given an indexed functor, we
derive the indexed functor coding its free monad:
\note{pwm: Whoa there. Maybe we should say something about IMonads in
general before we get to this point?}

@@

\[\stk{
\HExprAFreeD : \ATagIDesc{\Ty} \\
\HExprAFreeD \mapsto \bigRedBracket{
                 \begin{array}{l}
                   \EnumT{\red{[}\Econd{}{}{}\red{]}} \red{,} \\
                   \LAM{ty}
                   \bigRedBracket{
                   \begin{array}{l}
                   \DProd{\DVar{\Ebool}}{\DProd{\DVar{ty}}{\DVar{ty}}}
                   \end{array}
                   }
                 \end{array}
                 }
}\]

@@

\[\stk{
\IDescFreeD : \PI{\V{I}}{\Set} \IDesc{\Unit} \\
\begin{array}{@{}ll}
\IDescFreeD~I \mapsto \DSigma{}{} & (\EnumT \red{[} \DConst{},
                                              \DProd{}{},
                                              \DSigma{}{}, 
                                              \DPi{}{} \red{]}) \\
                                  & \bigRedBracket{\begin{array}{l}
                                        \DConst{\Set}               \\
                                        \DProd{\DVar{\Void}}{\DVar{\Void}}  \\
                                        \DSigma{\Set}{(\LAM{S} \DPi{S}{(\LAM{\_} \DVar{\Void})})} \\
                                        \DSigma{\Set}{(\LAM{S} \DPi{S}{(\LAM{\_} \DVar{\Void})})}
                                    \end{array}}
\end{array}
}\]

@@

\[\stk{
\IDescD : \PI{\V{I}}{\Set} \IDesc{\Unit} \\
\IDescD\: I \mapsto \FreeIMonad{\red{[}\IDescFreeD\red{]}}{\Unit}{I}\: I
}\]

### 2010-03-18 papers/icfp-2010-desc/paper_type_theory.tex
Desc@ICFP: 'bidirectional'
Desc@ICFP: Some minor fixes and notes

\begin{wstructure}
    <- Judgemental equality
        <- Presentation independent of particular implementation choice
        -> Model in Agda, intentional
        -> Used in Epigram, OTT
\end{wstructure}

Noticeably, we have adopted a judgemental presentation of
equality. Therefore, we are not tied to a particular implementation
choice. In particular, our system has been modelled in Agda, which
features an intentional equality. On the other hand, it is used in
Epigram, which equality has a slightly extensional
flavor~\cite{altenkirch:ott}. We expect users of fully extensional
systems to also find their way through this presentation.

@@

@@

In this section, we have developed a type propagation system based on
bidirectional type-checking. Using bidirectional type-checking as a
boilerplate scrapper is a well-known technique~\cite{who?} \note{I
  would need some citation of people using bidirectional presentation
  of their type theory}. In our case, we have shown how to instrument
bi-directionality to rationalise the expressivity of dependent
types. We have illustrated our approach with finite sets. We have
abstracted away the low-level presentation of finite sets, offering a
convenient syntax instead.

This example teaches us that we should not be afraid by codes, as soon
as type information is available. We have shown how to rationalise
this information in a formal presentation. Hence, we have proved that
programming with such objects is practical. As we introduce more codes
in our theory, we will show how to extend the framework we have
developed so far.

>>>>

\begin{wstructure}
    <- Judgemental equality
        <- Presentation independent of particular implementation choice
        -> Model in Agda, intensional
        -> Used in Epigram, OTT
\end{wstructure}

Noticeably, we have adopted a judgemental presentation of
equality. Therefore, we are not tied to a particular implementation
choice. In particular, our system has been modelled in Agda, which
features an intensional equality. On the other hand, it is used in
Epigram, which equality has a slightly extensional
flavor~\cite{altenkirch:ott}. We expect users of fully extensional
systems to also find their way through this presentation.

@@

\note{pwm: We should explain this new judgement form, specifically that
$\propag{}{}$ means `elaborates to'.}

@@

In this section, we have developed a type propagation system based on
bidirectional type-checking. Using bidirectional type-checking as a
boilerplate scrapper is a well-known technique~\cite{who?} \note{I
  would need some citation of people using bidirectional presentation
  of their type theory}. In our case, we have shown how to instrument
bidirectionality to rationalise the expressivity of dependent
types. We have illustrated our approach with finite sets. We have
abstracted away the low-level presentation of finite sets, offering a
convenient syntax instead.

This example teaches us that we should not be afraid by codes, as soon
as type information is available. We have shown how to rationalise
this information in a formal presentation. Hence, we have shown that
programming with such objects is practical. As we introduce more codes
in our theory, we will show how to extend the framework we have
developed so far.
